% ====================================================================
% Learned change of variables (pi_cov) -- finished-product framing.
% Placement: \subsubsection within Section~\ref{sec:open}, after the
% \subsubsection{Results} block (after the per-class paragraph).
% Companion edits to lines 145 and 269 and the removal of the line-360
% limitation are listed in the accompanying notes.
% ====================================================================
\subsubsection{Learning the change of variables}
\label{sec:picov}

The change-of-variables macroaction is driven by a learned policy
$\pi_{\mathrm{cov}}$: given an equation, $\pi_{\mathrm{cov}}$ generates the
substitution $x \mapsto \phi(x)$ that depresses it. The solver's main policy
learns \emph{when} to invoke a change of variables and how to compose it with
ordinary algebraic steps; $\pi_{\mathrm{cov}}$ supplies the substitution
itself.

\textbf{Task and data.}
$\pi_{\mathrm{cov}}$ maps an equation to its depression substitution. We train
it on dedicated datasets that instantiate the depressable forms of the four
families with coefficient symbols drawn, as permutations, from a fixed symbol
pool. No numeric coefficients appear, so $\pi_{\mathrm{cov}}$ cannot memorize
coefficient values and must learn the \emph{structural} depression rule. The
largest split, \texttt{cov\_large}, holds $774$ training and $192$ held-out
test equations; the test equations are symbol-permutations disjoint from
those seen in training.

\textbf{Training.}
The depressing substitution of every template is known in closed form, so
each training equation is paired with a target substitution, and
$\pi_{\mathrm{cov}}$ is trained by supervised learning to reproduce it. We
apply random symbol-renaming augmentation throughout, so $\pi_{\mathrm{cov}}$
learns the structural rule independently of the particular symbols a given
equation happens to use.

\textbf{Generating the substitution.}
$\pi_{\mathrm{cov}}$ emits the substitution's expression tree directly, as a
sequence of grammar productions in prefix order over the operators
$\{+,-,\times,\div,\log\}$, the variable $x$, the small integers
$\{2,3,4\}$, and a \textsc{copy} token naming a coefficient symbol of the
equation. For example, $x - b/(2a)$ is generated as the token sequence
\textsc{sub}\,\textsc{x}\,\textsc{div}\,\textsc{copy}\,\textsc{mul}\,%
\textsc{int2}\,\textsc{copy}; production arities make decoding
self-delimiting. The model thus chooses the substitution's structure itself
-- whether to subtract, what to divide by, which coefficients to use --
rather than filling a fixed template.

\textbf{Results.}
On the $192$ held-out test equations, $\pi_{\mathrm{cov}}$ produces the
correct substitution for \textbf{$100\%$}, identically across eight random
seeds, under \emph{deterministic} greedy decoding -- no search is required.
$\pi_{\mathrm{cov}}$ is evaluated here on symbol-permutations of the four
families; extending it to symbol pools and template families beyond those
seen in training is a natural direction for future work.
